# Useful utilities

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE, cache=TRUE)
```


## `bitr`: Biological Id TranslatoR


[clusterProfiler](https://www.bioconductor.org/packages/clusterProfiler) provides `bitr` and `bitr_kegg` for converting ID types. Both `bitr` and `bitr_kegg` support many species including model and many non-model organisms.

```{r}
x <- c("GPX3",  "GLRX",   "LBP",   "CRYAB", "DEFB1", "HCLS1",   "SOD2",   "HSPA2",
       "ORM1",  "IGFBP1", "PTHLH", "GPC3",  "IGFBP3","TOB1",    "MITF",   "NDRG1",
       "NR1H4", "FGFR3",  "PVR",   "IL6",   "PTPRM", "ERBB2",   "NID2",   "LAMB1",
       "COMP",  "PLS3",   "MCAM",  "SPP1",  "LAMC1", "COL4A2",  "COL4A1", "MYOC",
       "ANXA4", "TFPI2",  "CST6",  "SLPI",  "TIMP2", "CPM",     "GGT1",   "NNMT",
       "MAL",   "EEF1A2", "HGD",   "TCN2",  "CDA",   "PCCA",    "CRYM",   "PDXK",
       "STC1",  "WARS",  "HMOX1", "FXYD2", "RBP4",   "SLC6A12", "KDELR3", "ITM2B")
eg = bitr(x, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")
head(eg)
```

User should provides an annotation package, both _fromType_ and _toType_ can accept any types that supported.

User can use _keytypes_ to list all supporting types.

```{r}
library(org.Hs.eg.db)
keytypes(org.Hs.eg.db)
```

We can translate from one type to other types.
```{r}
ids <- bitr(x, fromType="SYMBOL", toType=c("UNIPROT", "ENSEMBL"), OrgDb="org.Hs.eg.db")
head(ids)
```

For GO analysis, user don't need to convert ID, all ID type provided by `OrgDb` can be used in `groupGO`, `enrichGO` and `gseGO` by specifying `keyType` parameter.

### `bitr_kegg`: converting biological IDs using KEGG API


```{r}
data(gcSample)
hg <- gcSample[[1]]
head(hg)

eg2np <- bitr_kegg(hg, fromType='kegg', toType='ncbi-proteinid', organism='hsa')
head(eg2np)
```

The ID type (both `fromType` & `toType`) should be one of 'kegg', 'ncbi-geneid', 'ncbi-proteinid' or 'uniprot'. The 'kegg' is the primary ID used in KEGG database. The data source of KEGG was from NCBI. A rule of thumb for the 'kegg' ID is `entrezgene` ID for eukaryote species and `Locus` ID for prokaryotes.

Many prokaryote species don't have entrezgene ID available. For example we can check the gene information of `ece:Z5100` in <http://www.genome.jp/dbget-bin/www_bget?ece:Z5100>, which have `NCBI-ProteinID` and `UnitProt` links in the `Other DBs` Entry, but not `NCBI-GeneID`.


If we try to convert `Z5100` to `ncbi-geneid`, `bitr_kegg` will throw error of `ncbi-geneid is not supported`.

```{r eval=FALSE}
bitr_kegg("Z5100", fromType="kegg", toType='ncbi-geneid', organism='ece')
```

```
## Error in KEGG_convert(fromType, toType, organism) :
## ncbi-geneid is not supported for ece ...
```

We can of course convert it to `ncbi-proteinid` and `uniprot`:

```{r}
bitr_kegg("Z5100", fromType="kegg", toType='ncbi-proteinid', organism='ece')
bitr_kegg("Z5100", fromType="kegg", toType='uniprot', organism='ece')
```

